Kalman tracking and parameter estimation of continuous gravitational waves with a pulsar timing array: inclusion of pulsar terms

### 2023-11-21 manuscript/manuscript_2/paper.tex
repro original plot, c.f. iota issues

In order to generate synthetic frequency time series $f_{\rm m}^{(1)}(t), \dots, f_{\rm m}^{(N)}(t)$, Equations \eqref{eq:frequency_evolution}--\eqref{eq:spinevol} are integrated numerically using a Runge-Kutta It$\hat{\text{o}}$ integrator implemented in the \texttt{sdeint} python package \footnote{\url{https://github.com/mattja/sdeint}}. This produces random realizations of $f_{\rm p}^{(n)}(t)$ for $1\leq n \leq N$. The measured $f_{\rm m}^{(n)}(t)$ are then produced via Equations \eqref{eq:measurement}--\eqref{eq:g_func_trig}. Equations \eqref{eq:frequency_evolution}--\eqref{eq:spinevol} and \eqref{eq:measurement}--\eqref{eq:g_func_trig} depend on the choice of static parameters $\boldsymbol{\theta'}$, as well as the choice of pulsar directions ${\boldsymbol{q}}^{(n)}$ and detector measurement noise covariance $\sigma_{\rm m}$ (Equation \eqref{eq:sigma_m_eqn}). We now describe how these values are chosen so as to generate astrophysically representative synthetic data. \newline

@@

\subsection{Parameter estimation} \label{sec:pe_and_ms}

The static GW source parameters $\boldsymbol{\theta}_{\rm gw}$ used for this mock analysis are selected to be astrophysically reasonable and representative. The injected components of $\boldsymbol{\theta}_{\rm gw}$ and $\boldsymbol{\theta}_{\rm psr}$ are summarised in the second column of Table \ref{tab:parameters_and_priors}. Different values of the static GW source parameters are explored in Section XX. \newline

The synthetic data generated by the above procedure is passed to the analysis scheme in Section \ref{sec:pe_and_ms}.

In this section, we calculate the posterior probability distribution for the static parameters $\boldsymbol{\theta} = \boldsymbol{\theta}_{\rm gw} \cup \boldsymbol{\theta}_{\rm psr}$ and compare it to the known, injected values. The procedure is undertaken for multiple realisations of the noise processes $\xi^{(n)}(t)$ and $\varepsilon^{(n)}(t)$. That is, given the injected static parameters $\boldsymbol{\theta}$ summarised in Table \ref{Table tab:parameters_and_priors}, multiple realisations of synthetic data are generated using the method described in Section \ref{sec:synthetic_data}, and for each realisation we independently estimate $\boldsymbol{\theta}$.The aims are (i) to demonstrate that the analysis scheme works (i.e.\ that it converges to a well-behaved, unimodal posterior for multiple noise realisations), (ii) to give a preliminary sense of its accuracy, and (iii) demonstrate that quantify the natural random dispersion in the one-dimensional posterior medians between noise realisations. We apply the Kalman filter in conjunction with nested sampling in order to infer the joint posterior distribution $p({\boldsymbol{\theta}} | {\boldsymbol{Y}})$. The specification of the priors on $\boldsymbol{\theta}$ is described in section \ref{sec:set_priors} and summarised in \ref{Table tab:parameters_and_priors}

>>>>

In order to generate synthetic frequency time series $ \boldsymbol{Y}(t) = \{f_{\rm m}^{(1)}(t), \dots, f_{\rm m}^{(N)}(t) \}$, Equations \eqref{eq:frequency_evolution}--\eqref{eq:spinevol} are integrated numerically using a Runge-Kutta It$\hat{\text{o}}$ integrator implemented in the \texttt{sdeint} python package \footnote{\url{https://github.com/mattja/sdeint}}. This produces random realizations of $f_{\rm p}^{(n)}(t)$ for $1\leq n \leq N$. The measured $f_{\rm m}^{(n)}(t)$ are then produced via Equations \eqref{eq:measurement}--\eqref{eq:g_func_trig}. Equations \eqref{eq:frequency_evolution}--\eqref{eq:spinevol} and \eqref{eq:measurement}--\eqref{eq:g_func_trig} depend on the choice of static parameters $\boldsymbol{\theta'}$, as well as the choice of pulsar directions ${\boldsymbol{q}}^{(n)}$ and detector measurement noise covariance $\sigma_{\rm m}$ (Equation \eqref{eq:sigma_m_eqn}). We now describe how $\boldsymbol{\theta}_{\rm psr}$, ${\boldsymbol{q}}^{(n)}$ and $\sigma_{\rm m}$ are chosen. This is equivalent to specifying the configuration of a synthetic PTA. In Section  \ref{sec:pe_and_ms} and Section \ref{sec:parameter_space} we describe how $\boldsymbol{\theta}_{\rm gw}$ chosen. This is equivalent to specifying the synthetic SMBHB source.  \newline

@@

\subsection{Parameter estimation} \label{sec:pe_and_ms} 
In this section, we calculate the joint posterior probability distribution $p({\boldsymbol{\theta}} | {\boldsymbol{Y}})$

for synthetic data $\boldsymbol{Y}$

The injected static GW source parameters $\boldsymbol{\theta}_{\rm gw}$ used for this mock analysis to generate $\boldsymbol{Y}$ are selected to be astrophysically reasonable and representative.

The static pulsar parameters $\boldsymbol{\theta'}_{\rm psr}$  are as specified in Section \ref{sec:synthetic_data}

The injected components of $\boldsymbol{\theta}_{\rm gw}$ and $\boldsymbol{\theta}_{\rm psr}$ are summarised in the second column of Table \ref{tab:parameters_and_priors}. The specification of the priors on $\boldsymbol{\theta}$ is described in Appendix \ref{sec:set_priors} and summarised in Table \ref{tab:parameters_and_priors}. Different values of the static GW source parameters are explored in Section \ref{sec:parameter_space}.

posterior probability distribution for the static parameters $\boldsymbol{\theta}$ and compare it to the known, injected values.

 The procedure is undertaken for multiple realisations of the noise processes $\xi^{(n)}(t)$ and $\varepsilon^{(n)}(t)$. That is, given the injected static parameters $\boldsymbol{\theta}$ multiple realisations of synthetic data are generated using the methods described in Section \ref{sec:synthetic_data}, and for each realisation we independently estimate $\boldsymbol{\theta}$. The aims are (i) to demonstrate that the analysis scheme works (i.e.\ that it converges to a well-behaved, unimodal posterior for multiple noise realisations), (ii) to give a preliminary sense of its accuracy, and (iii) demonstrate that quantify the natural random dispersion in the one-dimensional posterior medians between noise realisations.

We apply the Kalman filter in conjunction with nested sampling in order to infer the
